Exploring probability distributions for bivariate temporal granularities

### 2019-12-15 main.Rmd
choice of plots updated

Choice of plots are dictated by the statistical transformations and geometric objects used for the visualization. The basic plot choice is display of density. In Section \ref{sec:measured-variable}, it is dicussed that these densities could be estimated using Kernel density estimates or some descriptive statistics. The latter do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former. However, they avoid much clutter and help us to focus on some specific properties of the data. as well as on additional smoothing parameters and hence can vary for the same data set. In practice, the densities are estimated reasonably with at least 30 observations. However, even with large sample size, choosing too large a value for bandwidth can cause the estimate to oversmooth the data. The summaries based on descriptive statistics remain free of any tuning parameters or Kernels, which is sometimes more desirable for exploration.

@@

Quantile plots visually portray the quantiles of the distribution of sample data. These plots can display any quantiles instead of quartiles in a traditional boxplot. Some of the most implemented definitions of sample quantiles based on one or two order statistics can be found in @Hyndman1996-ty. These plots give more information than box plots and yet avoid clutter by enabling us to focus just on specific probabilities. When reviewing a boxplot, an outlier is defined as a data point that is located outside the fences (“whiskers”) of the boxplot (e.g. outside 1.5 times the interquartile range above the upper quartile and below the lower quartile). However, outliers are open to interpretation and not shown in an quantile plot.

@@

<!-- Description and lit review of density plots -->
 A Ridge line plot (sometimes called Joy plot) shows the distribution of a numeric value for several groups. Distribution can be represented using histograms or density plots, all aligned to the same horizontal scale and presented with a slight overlap. Like other density based displays, these plots allow us to see multimodality in the distribution. However, these plots can be obscuring when there is overlap of distribution for two or more categories of the y-axis. Also, with lot of categories, it is difficult to compare the height of the densities across categories.

>>>>

Choice of plots are dictated by the statistical transformations and geometric objects used for the visualization. The basic plot choice is display of density. In Section \ref{sec:measured-variable}, it is dicussed that these densities could be estimated using Kernel density estimates or some descriptive statistics. The latter do not allow us to see the shape, skewness, nature of the tail or multi-modality with so much clarity as the former. However, they avoid much clutter and help us to focus on some specific properties of the data. Kernel density estimates provide more detailed information about the distribution but depend on smoothing parameters and hence can vary for the same data set. In practice, the densities are estimated reasonably with at least 30 observations. However, even with large sample size, choosing too large a value for bandwidth can cause the estimate to oversmooth the data. The summaries based on descriptive statistics remain free of any tuning parameters or Kernels, which is sometimes more desirable for exploration.

@@

Quantile plots can display any quantiles instead of quartiles in a traditional boxplot. These plots give more information than box plots and yet avoid clutter by enabling us to focus just on specific probabilities. When reviewing a boxplot, an outlier is defined as a data point that is located outside the fences (“whiskers”) of the boxplot (e.g. outside 1.5 times the interquartile range above the upper quartile and below the lower quartile). However, outliers are open to interpretation and not shown in an quantile plot.

@@

<!-- Description and lit review of density plots -->
 A Ridge line plot (sometimes called Joy plot) shows the distribution for several groups. Distribution can be represented using density plots, all aligned to the same horizontal scale and presented with a slight overlap. Like other density based displays, these plots allow us to see multimodality in the distribution. However, these plots can be obscuring when there is overlap of distribution for two or more categories of the y-axis. Also, with lot of categories, it is difficult to compare the height of the densities across categories.
